---
title: "Final_Project"
author: "MJ"
date: "9/17/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# NOAA Storm data base exploration
## Synopsis: 
This script explores the most harmful weather events across the U.S.. To do this, the script first loads the required packages and the dataset. In a second step the script prepares the data for the analysis. That is, the events are defined as a factor and the data is subset into smaller dataframes. The data is subset by the consequences (fatalities, injuries, and damage) and the sum is calculated for each consequence and weather event. Subsequently, the data subsets are ordered and the top 10 weather events per consequence (i.e., fatalities, injuries, damage) are used as the input data for the plotting. The plotting generates bar plots with ggplot2 and the barplots for human damage (i.e., injuries and fatalities) are merged into one plot with the patchwork package.
```{r, echo = TRUE}
library(patchwork)
library(plyr)
library(ggplot2)
url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
download.file(url,"C:/Users/joerling/Dropbox/Sonstiges/Learning_Data_Science/Reproducible Research/Week 4/Project/StormData.csv")
df <- read.csv("C:/Users/joerling/Dropbox/Sonstiges/Learning_Data_Science/Reproducible Research/Week 4/Project/StormData.csv")
```

## Data processing

```{r, echo = TRUE}
df$EVTYPE <- as.factor(df$EVTYPE)

top_fatalities <- aggregate(FATALITIES ~ EVTYPE, df, FUN = sum)
top_fatalities <- arrange(top_fatalities,desc(top_fatalities$FATALITIES))
top_fatalities <- top_fatalities[1:10,]

top_injuries <- aggregate(INJURIES ~ EVTYPE, df, FUN = sum)
top_injuries <- arrange(top_injuries,desc(top_injuries$INJURIES))
top_injuries <- top_injuries[1:10,]

top_damage <- aggregate(PROPDMG ~ EVTYPE, df, FUN = sum)
top_damage <- arrange(top_damage,desc(top_damage$PROPDMG))
top_damage <- top_damage[1:10,]
```

## Results
```{r, echo = TRUE}
plot_fatalities <- ggplot(data = top_fatalities, aes(x = EVTYPE, y = FATALITIES))+
  geom_bar(stat = "identity") +
  labs(title = "Top 10 Fatalities", x = "Event Type", y = "Fatalities") +
  theme(axis.text.x = element_text(angle = 90))

plot_injuries <- ggplot(data = top_injuries, aes(x = EVTYPE, y = INJURIES))+
  geom_bar(stat = "identity") +
  labs(title = "Top 10 Injuries", x = "Event Type", y = "Injuries") +
  theme(axis.text.x = element_text(angle = 90))

plot_people_harm <- plot_fatalities + plot_injuries
plot_people_harm


plot_damage <- ggplot(data = top_damage, aes(x = EVTYPE, y = PROPDMG))+
  geom_bar(stat = "identity") +
  labs(title = "Top 10 Property Damage", x = "Event Type", y = "Damage in $") +
  theme(axis.text.x = element_text(angle = 90))
plot_damage
```

